---
title: "Vizualization"
author: "Ted Hart"
date: "August 5, 2014"
output: html_document
---

There are lot's of plotting options for `spocc`. At it's most basic, we can just plot our search results on a global map.


```{r global plot}
library(spocc)
bbox <- c(-80.79,38.64,-69.62,45.56)
### Searching for sugar maples in New England
out <- occ(query='Acer saccharum', from=c('inat','gbif'), geometry=bbox, limit = 100)
plot(out)
```

This is a great quick visualization, but what if we want to be able to zoom in on our plots?  There are a couple of choices.  Out of the box we can use `mapggplot` function.  It will automatically clip your plot to the extent of your search.

```{r ggplot}
mapggplot(out)
```

These are great out of the box plots, but we can also get more custom plots with `ggplot2` and the `maps` package.  We'll plot just the states in our bounding box and overlay our species data.

```{r ggplot custom}
library(maps)

states <- map_data("state")
n_east <- subset(states, region %in% c("vermont","new hampshire","new york","connecticut","pennsylvania","new jersey","rhode island","massachusetts","maryland"  ) )
ne_map <- ggplot() +geom_polygon(data = n_east,aes(x=long,y=lat,group = group), colour="white",fill="grey60") + theme_bw()
ne_map
### Add points
ne_map + geom_point(data = occ2df(out),aes(x=longitude,y=latitude))
```

With the base `ggplot` map we can easily make more complicated maps.

```{r ggplot custom 2}
# Redo search adding extra species
out <- occ(query=c('Acer saccharum','Acer rubrum'), from=c('gbif'), geometry=bbox, limit = 150)

ne_map + geom_point(data = occ2df(out),aes(x=longitude,y=latitude, group= name, colour = name, size = 1.5)) + scale_size(guide=F) + scale_colour_manual(values=c("red","blue"))

```

Another option in `spocc` is to make maps with some nice javascript maps as well.

```{r leaflet}
out <- occ(query=c('Acer saccharum'), from=c('gbif'), geometry=bbox, limit = 150)
out <- occ2df(out)
mapleaflet(data = out, dest = ".")

```

We can also make maps with geoJSON and render them on github (if you have a github account)

```{r geojson}
out <- occ(query=c('Acer saccharum','Acer rubrum'), from='gbif',gbifopts = list(hasCoordinate=TRUE))
out <- fixnames(out)
out <- occ2df(out)
mapgist(data=out, color=c("#00ff00","#ff00ff"))

```


